---
title: "HW 1"
author: "Shivani Patel"
date: "2024-08-28"
output: html_document
---

# 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called ‘Class 1 Survey Fall 2024_di.csv’), calling the R dataframe that you create C1survey.

```{r}
Class_fall <-
 read_csv(
   "https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv"
   )
```

# 2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.


```{r}
nrow(Class_fall)
ncol(Class_fall)
```

# 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.


```{r}
names(Class_fall)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan",
"fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby",
"program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length",
"top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state",
"city", "highest_educ_level")
# create list of renamed variables in the object varlist
colnames(Class_fall)

```
# 4.  Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.

```{r}

x <- sapply(Class_fall, class) 

#to calculate number of each observation
x_counts <- table(x)

 # display counts
x_counts
```




# 5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics


```{r}
summary(Class_fall$bday)
summary(Class_fall$bmonth)
```
```{r}
unique(Class_fall$bday)
unique(Class_fall$bmonth)
```
```{r}
Class_fall$bday<-as.numeric(Class_fall$bday)
Class_fall$bmonth<-as.numeric(Class_fall$bmonth)

# check if variables are now structurally numeric

str(Class_fall$bday)
str(Class_fall$bmonth)
```

```{r}

#to see frequency of the unusual or missing values

table(Class_fall$bday, useNA = "ifany")
table(Class_fall$bmonth, useNA = "ifany")
```
```{r}

```
 

# 6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

#a) Creating a new variable 'bseason' based on birth month

```{r}
Class_fall <- Class_fall 
  mutate(bseason = 
           case_when(bmonth %in% c(12, 1, 2) ~ "winter", #Months 12, 1, 2 -> Winter
                     bmonth %in% c(3, 4, 5) ~ "spring", #Months 3, 4, 5 -> Spring
                     bmonth %in% c(6, 7, 8) ~ "summer", #Months 6, 7, 8 -> Summer
                     bmonth %in% c(9, 10, 11) ~ "fall", #Months 9, 10, 11 -> Fall
                     ))
```


#a) Creating a seasons table : seasons as columns and birth month as row
```{r}
seasons_table <- table(Class_fall$bmonth, Class_fall$bseason, useNA = "ifany")
seasons_table
```










